---
title: "Integrated Risk Management Model"
author: "Amanda"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
    theme: cosmo
    code_folding: hide
    df_print: paged
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```
#1. Market Risk Engine – Parametric & Monte Carlo VaR
```{r eruptions, echo=FALSE}


library(quantmod)
library(PerformanceAnalytics)
library(MASS)

# Get sample data (e.g., AAPL, MSFT, JPM, GOOGL)
tickers <- c("AAPL", "MSFT", "GOOGL", "JPM")
getSymbols(tickers, from = "2020-01-01")

# Get adjusted close prices
prices <- do.call(merge, lapply(tickers, function(t) Cl(get(t))))
returns <- na.omit(Return.calculate(prices))

# Assume equal weights
weights <- rep(1/ncol(returns), ncol(returns))

# Calculate portfolio return stats
mu <- colMeans(returns)
cov_matrix <- cov(returns)
port_sd <- sqrt(t(weights) %*% cov_matrix %*% weights)

# Parametric VaR (95%)
var_95 <- qnorm(0.05, mean = 0, sd = port_sd)
es_95 <- -port_sd * dnorm(qnorm(0.05)) / 0.05

cat("Portfolio Std Dev:", round(port_sd, 6), "\n")
paste("Parametric VaR (95%):", round(var_95, 6))
cat("Expected Shortfall (95%):", round(es_95, 6), "\n")

# Simulate portfolio returns (Monte Carlo)
set.seed(123)
n <- 10000
sim_returns <- mvrnorm(n, mu = mu, Sigma = cov_matrix)
portfolio_sim <- sim_returns %*% weights

# Calculate simulated VaR and ES
mc_var_95 <- quantile(portfolio_sim, 0.05)
mc_es_95 <- mean(portfolio_sim[portfolio_sim <= mc_var_95])

cat("Monte Carlo VaR (95%):", round(mc_var_95, 6), "\n")
cat("Monte Carlo ES (95%):", round(mc_es_95, 6), "\n")

```
#2. Credit risk 
```{r}
#Credit risk 
set.seed(42)

# Define credit risk parameters
n_sim <- 10000          # Number of simulations
pd <- 0.05              # 5% chance of default
lgd <- 0.6              # 60% loss if default happens
ead <- 100000           # Exposure at default (e.g., bond value)

# Simulate credit losses
credit_losses <- rbinom(n_sim, 1, pd) * lgd * ead

# Calculate risk metrics
credit_var_95 <- quantile(credit_losses, 0.95)
credit_es_95 <- mean(credit_losses[credit_losses > credit_var_95])

cat("Credit VaR (95%):", round(credit_var_95, 2), "\n")
cat("Credit Expected Shortfall (95%):", round(credit_es_95, 2), "\n")

# Visualization of Credit Loss Distribution
hist(credit_losses, breaks = 30, col = 'tomato', main = "Simulated Credit Loss Distribution", xlab = "Loss Amount")
abline(v = credit_var_95, col = "blue", lwd = 2, lty = 2)


```
#3. Integrated Risk-Model
```{r}
# Integrated Risk

market_sim <- portfolio_sim[1:10000]
credit_sim <- credit_losses[1:10000]

# Total integrated losses
total_loss <- -market_sim + credit_sim  # note: market loss = -return

# Calculate integrated risk metrics
integrated_var_95 <- quantile(total_loss, 0.95)
integrated_es_95 <- mean(total_loss[total_loss > integrated_var_95])

cat("Integrated VaR (95%):", round(integrated_var_95, 2), "\n")
cat("Integrated Expected Shortfall (95%):", round(integrated_es_95, 2), "\n")

```
#4. Scenario-based simulation
```{r}
# Stress test: 10% drop in equity, PD spikes to 25%
set.seed(123)
shock_equity <- market_sim * 0.9    # simulate 10% drop
shock_credit <- rbinom(length(shock_equity), 1, 0.25) * lgd * ead

stress_loss <- -shock_equity + shock_credit

# Stress VaR/ES
stress_var_95 <- quantile(stress_loss, 0.95)
stress_es_95 <- mean(stress_loss[stress_loss > stress_var_95])

cat("Stress Test VaR (95%):", round(stress_var_95, 2), "\n")
cat("Stress Test ES (95%):", round(stress_es_95, 2), "\n")
```
#5. VaR Breach Counter
```{r}
library(quantmod)
library(PerformanceAnalytics)

library(quantmod)
library(PerformanceAnalytics)

# Use your same tickers
tickers <- c("AAPL", "MSFT", "GOOGL", "JPM")
getSymbols(tickers, from = "2020-01-01")

# Adjusted prices and returns
prices <- do.call(merge, lapply(tickers, function(t) Cl(get(t))))
returns <- na.omit(Return.calculate(prices))
weights <- rep(1/ncol(returns), ncol(returns))
portfolio_returns <- Return.portfolio(returns, weights = weights)

# Rolling 1-year window (252 trading days)
window <- 252
rolling_var <- rollapply(portfolio_returns, width = window, FUN = function(x) {
  sd(x) * qnorm(0.05)
}, by.column = FALSE, align = "right")

# Backtesting comparison
actual_returns <- portfolio_returns[(window+1):NROW(portfolio_returns)]
predicted_var <- rolling_var

# Align lengths safely
min_len <- min(length(actual_returns), length(predicted_var))
actual_returns <- actual_returns[1:min_len]
predicted_var <- predicted_var[1:min_len]

# Remove NA or infinite values
valid_idx <- which(is.finite(actual_returns) & is.finite(predicted_var))
actual_returns <- actual_returns[valid_idx]
predicted_var <- predicted_var[valid_idx]

summary(actual_returns)
summary(predicted_var)
any(!is.finite(actual_returns))   # should return FALSE
any(!is.finite(predicted_var))    # should return FALSE

# Count breaches
breaches <- actual_returns < predicted_var
breach_count <- sum(breaches)
total <- length(breaches)

cat("Total Observations:", total, "\n")
cat("Number of VaR Breaches:", breach_count, "\n")
cat("Breach Rate:", round(breach_count / total, 4), "\n")



library(GAS)
library(rugarch)

# Kupiec test
kupiec_test <- VaRTest(alpha = 0.95, actual = as.numeric(actual_returns), VaR = as.numeric(predicted_var))
print(kupiec_test)

# Additional backtest
BacktestVaR(
  data = as.numeric(actual_returns), 
  VaR = as.numeric(predicted_var), 
  alpha = 0.95, 
  Lags = 4
)


# Plot results
plot(actual_returns, type = "l", main = "VaR Backtest: Actual vs VaR", col = "black")
lines(predicted_var, col = "red")
points(which(breaches), actual_returns[breaches], col = "blue", pch = 16)
legend("bottomleft", legend = c("Actual Returns", "VaR (95%)", "Breaches"), 
       col = c("black", "red", "blue"), lty = c(1, 1, NA), pch = c(NA, NA, 16))


```
#6. Machine Learning-Based PD Estimator
```{r}
library(randomForest)
library(caret)

# Simulated credit dataset
set.seed(123)
n <- 1000
credit_data <- data.frame(
  leverage = runif(n, 0.1, 1.5),            # Debt/Assets
  interest_coverage = runif(n, 0.01, 0.2),  # EBIT/Interest
  rating_score = sample(1:5, n, replace = TRUE),  # Categorical proxy
  default = as.factor(rbinom(n, 1, 0.1))               # Binary outcome
)

# Split data
train_idx <- sample(1:n, size = 0.8 * n)
train_data <- credit_data[train_idx, ]
test_data <- credit_data[-train_idx, ]

# Train random forest
rf_model <- randomForest(default ~ ., data = train_data, ntree = 100)


# Predict on test set
test_probs <- predict(rf_model, newdata = test_data, type = "prob")[, 2]

# Evaluate model
conf_matrix <- confusionMatrix(as.factor(predict(rf_model, newdata = test_data)), as.factor(test_data$default))
print(conf_matrix)

```

```{r}
# Use the predicted PDs to simulate credit losses
sim_pd <- test_probs
lgd <- 0.6
ead <- 100000
sim_credit_loss <- rbinom(length(sim_pd), 1, sim_pd) * lgd * ead

# Calculate risk metrics
ml_credit_var <- quantile(sim_credit_loss, 0.95)
ml_credit_es <- mean(sim_credit_loss[sim_credit_loss > ml_credit_var])

cat("ML-Based Credit VaR (95%):", round(ml_credit_var, 2), "\n")
cat("ML-Based Credit ES (95%):", round(ml_credit_es, 2), "\n")

```


# 7. Final Report Summary

library(knitr)

summary_table <- data.frame(
  Metric = c(
    "Monte Carlo VaR (95%)",
    "Monte Carlo Expected Shortfall (95%)",
    "Credit VaR (95%)",
    "Credit Expected Shortfall (95%)",
    "Integrated VaR (95%)",
    "Integrated Expected Shortfall (95%)",
    "Stress Test VaR (95%)",
    "Stress Test Expected Shortfall (95%)",
    "ML-Based Credit VaR (95%)",
    "ML-Based Credit Expected Shortfall (95%)"
  ),
  Value = round(c(
    mc_var_95, mc_es_95,
    credit_var_95, credit_es_95,
    integrated_var_95, integrated_es_95,
    stress_var_95, stress_es_95,
    ml_credit_var, ml_credit_es
  ), 2)
)

Knit::kable(summary_table, caption = "Summary of Risk Metrics")

```
